---
title: "ihs4_cleaning"
author: "Kevin Tang"
date: "3/3/2020"
output: html_document
---

```{r}
#--- Load libraries - reminder to always load tidyverse last
library(epiDisplay)
library(plyr)
library(magrittr)
library(foreign)
library(psych)
library(readxl)
library(tidyverse)

#--- Change to show no scientific notation & round to 3 decimal places
options(scipen = 10, digits=3) 

#--- Set the plot theme to come out in black and white
theme_set(theme_bw())

#--- Read in the file
ihs4 <- read.csv("/Users/kevintang/Library/Mobile Documents/com~apple~CloudDocs/Documents/LSHTM/phd epidemiology/2019 2020/data/hces/mwi/ihs2016/hh_mod_g1.csv")
fcode <- read.csv("/Users/kevintang/Library/Mobile Documents/com~apple~CloudDocs/Documents/LSHTM/phd epidemiology/2019 2020/analysis/mwi_ihs4/mwi_item_code.csv")
ucode <- read.csv("/Users/kevintang/Library/Mobile Documents/com~apple~CloudDocs/Documents/LSHTM/phd epidemiology/2019 2020/analysis/mwi_ihs4/mwi_unit_code.csv")
```

>CHANGE VARIABLE NAMES AND DATA CLASS

Codebook (https://microdata.worldbank.org/index.php/catalog/2939/datafile/F192)

id =            case_id              purc_quant =    hh_g04a              gift_quant =    hh_g04a
consYN =        hh_g01               purc_unitA =    hh_g04b              gift_unitA =    hh_g04b
item_code =     hh_g02               purc_unit =     hh_g04b_label        gift_unit =     hh_g04b_label
item_oth =      hh_g01_oth           purc_unit_oth = hh_g04b_oth          gift_unit_oth = hh_g04b_oth
                                     purc_unit_size= hh_g04c              gift_unit_size= hh_g04c
                                     purc_price =    hh_g05
             
cons_quant =    hh_g03a              prod_quant =    hh_g06a
cons_unitA =    hh_g03b              prod_unitA =    hh_g06b
cons_unit =     hh_g03b_label        prod_unit =     hh_g06b_label
cons_unit_oth = hh_g03b_oth          prod_unit_oth = hh_g06b_oth
cons_unit_size= hh_g03c              prod_unit_size= hh_g06c

```{r}

#Summary info
names(ihs4)[names(ihs4) == 'case_id'] <- 'id'
names(ihs4)[names(ihs4) == 'hh_g01'] <- 'consYN'
names(ihs4)[names(ihs4) == 'hh_g02'] <- 'item_code'
names(ihs4)[names(ihs4) == 'hh_g01_oth'] <- 'item_oth'

ihs4$id <- as.character(ihs4$id)
ihs4$consYN <- as.factor(ihs4$consYN)
ihs4$item_code <- as.factor(ihs4$item_code)
ihs4$item_oth <- as.factor(ihs4$item_oth)

#G03: Quantity and units consumed of each food 
names(ihs4)[names(ihs4) == 'hh_g03a'] <- 'cons_quant'
names(ihs4)[names(ihs4) == 'hh_g03b'] <- 'cons_unitA'
names(ihs4)[names(ihs4) == 'hh_g03b_label'] <- 'cons_unit'
names(ihs4)[names(ihs4) == 'hh_g03b_oth'] <- 'cons_unit_oth'
names(ihs4)[names(ihs4) == 'hh_g03c'] <- 'cons_unit_size'

ihs4$cons_quant <- as.numeric(ihs4$cons_quant)
ihs4$cons_unitA <- as.factor(ihs4$cons_unitA)
ihs4$cons_unit <- as.factor(ihs4$cons_unit)
ihs4$cons_unit_oth <- as.factor(ihs4$cons_unit_oth)
ihs4$cons_unit_size <- as.factor(ihs4$cons_unit_size)


#G04: Quantity and units purchased of each food 
names(ihs4)[names(ihs4) == 'hh_g04a'] <- 'purc_quant'
names(ihs4)[names(ihs4) == 'hh_g04b'] <- 'purc_unitA'
names(ihs4)[names(ihs4) == 'hh_g04b_label'] <- 'purc_unit'
names(ihs4)[names(ihs4) == 'hh_g04b_oth'] <- 'purc_unit_oth'
names(ihs4)[names(ihs4) == 'hh_g04c'] <- 'purc_unit_size'
names(ihs4)[names(ihs4) == 'hh_g05'] <- 'purc_price'

ihs4$purc_quant <- as.numeric(ihs4$purc_quant)
ihs4$purc_unitA <- as.factor(ihs4$purc_unitA)
ihs4$purc_unit <- as.factor(ihs4$purc_unit)
ihs4$purc_unit_oth <- as.factor(ihs4$purc_unit_oth)
ihs4$purc_unit_size <- as.factor(ihs4$purc_unit_size)
ihs4$purc_price <- as.numeric(ihs4$purc_price)

#G06: Quantity and units home produced of each food 
names(ihs4)[names(ihs4) == 'hh_g06a'] <- 'prod_quant'
names(ihs4)[names(ihs4) == 'hh_g06b'] <- 'prod_unitA'
names(ihs4)[names(ihs4) == 'hh_g06b_label'] <- 'prod_unit'
names(ihs4)[names(ihs4) == 'hh_g06b_oth'] <- 'prod_unit_oth'
names(ihs4)[names(ihs4) == 'hh_g06c'] <- 'prod_unit_size'

ihs4$prod_quant <- as.numeric(ihs4$prod_quant)
ihs4$prod_unitA <- as.factor(ihs4$prod_unitA)
ihs4$prod_unit <- as.factor(ihs4$prod_unit)
ihs4$prod_unit_oth <-as.factor(ihs4$prod_unit_oth)
ihs4$prod_unit_size <- as.factor(ihs4$prod_unit_size)

#G07: Quantity and units gifted/other source of each food 
names(ihs4)[names(ihs4) == 'hh_g07a'] <- 'gift_quant'
names(ihs4)[names(ihs4) == 'hh_g07b'] <- 'gift_unitA'
names(ihs4)[names(ihs4) == 'hh_g07b_label'] <- 'gift_unit'
names(ihs4)[names(ihs4) == 'hh_g07b_oth'] <- 'gift_unit_oth'
names(ihs4)[names(ihs4) == 'hh_g07c'] <- 'gift_unit_size'

ihs4$gift_quant <- as.numeric(ihs4$gift_quant)
ihs4$gift_unitA <- as.factor(ihs4$gift_unitA)
ihs4$gift_unit <- as.factor(ihs4$gift_unit)
ihs4$gift_unit_oth <- as.factor(ihs4$gift_unit_oth)
ihs4$gift_unit_size <- as.factor(ihs4$gift_unit_size)
```

>MERGE IN UNIT AND ITEM LABELS

```{r}
ihs4 <- merge(x=ihs4, y=fcode, by.x='item_code', by.y='code', fill=-9999, all.x = TRUE)
ihs4 <- merge(x=ihs4, y=ucode, by.x='cons_unit', by.y='code', fill=-9999, all.x = TRUE)
ihs4 <- ihs4 %>% arrange(cons_unit) %>% arrange(item_code) %>% arrange(HHID)
ihs4 <- ihs4 %>% select(id:consYN, item_code, item, item_oth:cons_unitA, cons_unit, unit, cons_unit_oth:gift_unit_size)
```


>CLEANING UNITS LABELLED AS "OTHER"

There are a number of item units recorded as "23" but described as existing ucodes in our codebook.
We need to change these units from "23" to their appropriate units as in this analysis any item without a standard unit will be excluded. 
```{r}
ihs4 %>% filter(cons_unitA==23)
ihs4 %>% filter(cons_unit_oth!= "") %>% count(cons_unit_oth) %>% arrange(desc(n)) #which units are listed as "other"
ihs4 %>% filter(cons_unitA==23) %>% count(cons_unit_oth) %>% arrange(desc(n)) #which units are listed as "other" have not been converted to standard unit

ihs4 %>% filter(., cons_unit_oth=="HEAP") %>% select(item, item_code, cons_quant, cons_unitA, cons_unit, cons_unit_oth) %>% arrange(item)
ihs4 %>% filter(item_code=="803") %>% select(item, item_code, cons_quant, cons_unitA, cons_unit, cons_unit_oth) %>% arrange(desc(cons_unitA))

```

```{r}
ihs4[ihs4$cons_unit_oth == "HEAP", "cons_unitA"] <- 10
ihs4[ihs4$cons_unit_oth == "GRAMS", "cons_unitA"] <- 18
ihs4[ihs4$cons_unit_oth == "KILOGRAM", "cons_unitA"] <- 1
ihs4[ihs4$cons_unit_oth == "PAIL", "cons_unitA"] <- 4
ihs4[ihs4$cons_unit_oth == "TUBE /SACHET", "cons_unitA"] <- 22
ihs4[ihs4$cons_unit_oth == "SATCHET (MEDIUM)", "cons_unitA"] <- 42
ihs4[ihs4$cons_unit_oth == "HEAP (SMALL)", "cons_unitA"] <- 10
ihs4[ihs4$cons_unit_oth == "BUNDLE", "cons_unitA"] <- 8
ihs4[ihs4$cons_unit_oth == "HEAP (LARGE)", "cons_unitA"] <- "10C"
ihs4[ihs4$cons_unit_oth == "TABLESPOON", "cons_unitA"] <- 50
ihs4[ihs4$cons_unit_oth == "5L BUCKET", "cons_unitA"] <- 26
ihs4[ihs4$cons_unit_oth == "TUBE (MEDIUM)", "cons_unitA"] <- 42
ihs4[ihs4$cons_unit_oth == "PAIL (SMALL)", "cons_unitA"] <- 4
ihs4[ihs4$cons_unit_oth == "TEASPOON", "cons_unitA"] <- 20
ihs4[ihs4$cons_unit_oth == "HEAP (MEDIUM)", "cons_unitA"] <- 10
ihs4[ihs4$cons_unit_oth == "PACKET 100 GRAM", "cons_unitA"] <- 34
ihs4[ihs4$cons_unit_oth == "PAIL (LARGE)", "cons_unitA"] <- 4
ihs4[ihs4$cons_unit_oth == "TUBE (LARGE)", "cons_unitA"] <- 43
ihs4[ihs4$cons_unit_oth == "PACKET 150 GRAM", "cons_unitA"] <- 34
ihs4[ihs4$cons_unit_oth == "PIECE (SMALL)", "cons_unitA"] <- "9A"
ihs4[ihs4$cons_unit_oth == "PIECES", "cons_unitA"] <- 9
ihs4[ihs4$cons_unit_oth == "NO. 12 PLATE", "cons_unitA"] <- 7
ihs4[ihs4$cons_unit_oth == "SATCHET (LARGE)", "cons_unitA"] <- 43
ihs4[ihs4$cons_unit_oth == "PIECE MEDIUM", "cons_unitA"] <- "9B"
ihs4[ihs4$cons_unit_oth == "PIECE", "cons_unitA"] <- 9
ihs4[ihs4$cons_unit_oth == "NO. 10 PLATE", "cons_unitA"] <- 6
ihs4[ihs4$cons_unit_oth == "5 LITRES", "cons_unitA"] <- 26
ihs4[ihs4$cons_unit_oth == "PAIL (MEDIUM)", "cons_unitA"] <- "4B"
ihs4[ihs4$cons_unit_oth == "PIECE (LARGE)", "cons_unitA"] <- "9C"
ihs4[ihs4$cons_unit_oth == "MANGO", "cons_unitA"] <- 9
ihs4[ihs4$cons_unit_oth == "PACKET (SMALL)", "cons_unitA"] <- 54
ihs4[ihs4$cons_unit_oth == "SATCHET (250 GRAM)", "cons_unitA"] <- 65
ihs4[ihs4$cons_unit_oth == "SATCHET (20 GRAM)", "cons_unitA"] <- 41
ihs4[ihs4$cons_unit_oth == "TIN", "cons_unitA"] <- 72
ihs4[ihs4$cons_unit_oth == "PIECE (MEDIUM)", "cons_unitA"] <- "9B"
ihs4[ihs4$cons_unit_oth == "PACKET 250 GRAM", "cons_unitA"] <- 65
ihs4[ihs4$cons_unit_oth == "PACKETS", "cons_unitA"] <- 60
ihs4[ihs4$cons_unit_oth == "TIN (500G)", "cons_unitA"] <- 73
ihs4[ihs4$cons_unit_oth == "SACHET", "cons_unitA"] <- 22
ihs4[ihs4$cons_unit_oth == "SATCHET (MEDIUM)", "cons_unitA"] <- 22
ihs4[ihs4$cons_unit_oth == "SATCHET (500 GRAM)", "cons_unitA"] <- 22
ihs4[ihs4$cons_unit_oth == "SATCHET (SMALL)", "cons_unitA"] <- 22
ihs4[ihs4$cons_unit_oth == "SATCHET (150 GRAM)", "cons_unitA"] <- 43
ihs4[ihs4$cons_unit_oth == "BOTTLE" & ihs4$item_code == "803", "cons_unitA"] <- 15

ihs4[ihs4$cons_unit_oth == "HEAP", "cons_unit"] <- 10
ihs4[ihs4$cons_unit_oth == "GRAMS", "cons_unit"] <- 18
ihs4[ihs4$cons_unit_oth == "KILOGRAM", "cons_unit"] <- 1
ihs4[ihs4$cons_unit_oth == "PAIL", "cons_unit"] <- 4
ihs4[ihs4$cons_unit_oth == "TUBE /SACHET", "cons_unit"] <- 22
ihs4[ihs4$cons_unit_oth == "SATCHET (MEDIUM)", "cons_unit"] <- 42
ihs4[ihs4$cons_unit_oth == "HEAP (SMALL)", "cons_unit"] <- 10
ihs4[ihs4$cons_unit_oth == "BUNDLE", "cons_unit"] <- 8
ihs4[ihs4$cons_unit_oth == "HEAP (LARGE)", "cons_unit"] <- 10
ihs4[ihs4$cons_unit_oth == "TABLESPOON", "cons_unit"] <- 50
ihs4[ihs4$cons_unit_oth == "5L BUCKET", "cons_unit"] <- 26
ihs4[ihs4$cons_unit_oth == "TUBE (MEDIUM)", "cons_unit"] <- 42
ihs4[ihs4$cons_unit_oth == "PAIL (SMALL)", "cons_unit"] <- 4
ihs4[ihs4$cons_unit_oth == "TEASPOON", "cons_unit"] <- 20
ihs4[ihs4$cons_unit_oth == "HEAP (MEDIUM)", "cons_unit"] <- 10
ihs4[ihs4$cons_unit_oth == "PACKET 100 GRAM", "cons_unit"] <- 34
ihs4[ihs4$cons_unit_oth == "PAIL (LARGE)", "cons_unit"] <- 4
ihs4[ihs4$cons_unit_oth == "TUBE (LARGE)", "cons_unit"] <- 43
ihs4[ihs4$cons_unit_oth == "PACKET 150 GRAM", "cons_unit"] <- 34
ihs4[ihs4$cons_unit_oth == "PIECE (SMALL)", "cons_unit"] <- 9
ihs4[ihs4$cons_unit_oth == "PIECES", "cons_unit"] <- 9
ihs4[ihs4$cons_unit_oth == "NO. 12 PLATE", "cons_unit"] <- 7
ihs4[ihs4$cons_unit_oth == "SATCHET (LARGE)", "cons_unit"] <- 43
ihs4[ihs4$cons_unit_oth == "PIECE MEDIUM", "cons_unit"] <- 9
ihs4[ihs4$cons_unit_oth == "PIECE", "cons_unit"] <- 9
ihs4[ihs4$cons_unit_oth == "NO. 10 PLATE", "cons_unit"] <- 6
ihs4[ihs4$cons_unit_oth == "5 LITRES", "cons_unit"] <- 26
ihs4[ihs4$cons_unit_oth == "PAIL (MEDIUM)", "cons_unit"] <- 4
ihs4[ihs4$cons_unit_oth == "PIECE (LARGE)", "cons_unit"] <- 9
ihs4[ihs4$cons_unit_oth == "MANGO", "cons_unit"] <- 9
ihs4[ihs4$cons_unit_oth == "PACKET (SMALL)", "cons_unit"] <- 54
ihs4[ihs4$cons_unit_oth == "SATCHET (250 GRAM)", "cons_unit"] <- 65
ihs4[ihs4$cons_unit_oth == "SATCHET (20 GRAM)", "cons_unit"] <- 41
ihs4[ihs4$cons_unit_oth == "TIN", "cons_unit"] <- 72
ihs4[ihs4$cons_unit_oth == "PIECE (MEDIUM)", "cons_unit"] <- 9
ihs4[ihs4$cons_unit_oth == "PACKET 250 GRAM", "cons_unit"] <- 65
ihs4[ihs4$cons_unit_oth == "PACKETS", "cons_unit"] <- 60
ihs4[ihs4$cons_unit_oth == "TIN (500G)", "cons_unit"] <- 73
ihs4[ihs4$cons_unit_oth == "SACHET", "cons_unit"] <- 22
ihs4[ihs4$cons_unit_oth == "SATCHET (MEDIUM)", "cons_unit"] <- 22
ihs4[ihs4$cons_unit_oth == "SATCHET (500 GRAM)", "cons_unit"] <- 22
ihs4[ihs4$cons_unit_oth == "SATCHET (SMALL)", "cons_unit"] <- 22
ihs4[ihs4$cons_unit_oth == "SATCHET (150 GRAM)", "cons_unit"] <- 43
ihs4[ihs4$cons_unit_oth == "BOTTLE" & ihs4$item_code == "803", "cons_unit"] <- 15
```

>CLEANING UNITS LABELLED AS "OTHER"

```{r}
item_other <- ihs4 %>% count(item_oth) %>% arrange(desc(n))

ihs4 %>% filter(item_oth=="BIRD")
```

```{r}
ihs4$item_code[ihs4$item_oth=='BAKED AFRICAN CAKE'] <- 836
ihs4$item_code[ihs4$item_oth=='GRAIN MEAL'] <- 101
ihs4$item_code[ihs4$item_oth=='GRANMILL'] <- 101
ihs4$item_code[ihs4$item_oth=='MANDASI'] <- 827
ihs4$item_code[ihs4$item_oth=='PORK'] <- 506
ihs4$item_code[ihs4$item_oth=='PEANUT'] <- 304  
ihs4$item_code[ihs4$item_oth=='PIGEON PEAS'] <- 303  
ihs4$item_code[ihs4$item_oth=='CUCUMBER'] <- 409  
ihs4$item_code[ihs4$item_oth=='SMALL FISH'] <- 5023   
ihs4$item_code[ihs4$item_oth=='SODA'] <- 907 
ihs4$item_code[ihs4$item_oth=='FRIED MIXTURE OF MAIZE FLOUR/BANANA AND SODA'] <- 827 
ihs4$item_code[ihs4$item_oth=='MIXTURE OF MAIZE FLOUR/BANANA/SUGAR'] <- 827 
ihs4$item_code[ihs4$item_oth=='IRISH POTATOES AND FLOUR'] <- 821
ihs4$item_code[ihs4$item_oth=='SCONES'] <- 112 
ihs4$item_code[ihs4$item_oth=='WINE'] <- 914
ihs4$item_code[ihs4$item_oth=='TEA'] <- 901
ihs4$item_code[ihs4$item_oth=='BIRD'] <- 509
ihs4$item_code[ihs4$item_oth=='DUCKMEAT'] <- 509
ihs4$item_code[ihs4$item_oth=='COW PEAS'] <- 308
ihs4$item_code[ihs4$item_oth=='BONONGWE'] <- 403
ihs4$item_code[ihs4$item_oth=='CASSAVA LEAVES'] <- 406
ihs4$item_code[ihs4$item_oth=='CHIKONDAMOYO'] <- 836
ihs4$item_code[ihs4$item_oth=='CHISOSOANDBONONGWE'] <- 403
ihs4$item_code[ihs4$item_oth=='FRESHBEANS'] <- 302
ihs4$item_code[ihs4$item_oth=='GROUND AND COOKED BEANS'] <- 302
ihs4$item_code[ihs4$item_oth=='MAIZEMEAL'] <- 101
ihs4$item_code[ihs4$item_oth=='SOYA'] <- 314

```

> MISC. CLEANING FUNCTIONS

Create unique ID variable
```{r}
ihs4$measure_id <- paste0(as.character(ihs4$item_code),"_", as.character(ihs4$cons_unitA))
ihs4$measure_id[is.na(ihs4$cons_unit)] <- NA

write.csv(ihs4, "/Users/kevintang/Library/Mobile Documents/com~apple~CloudDocs/Documents/LSHTM/phd epidemiology/2019 2020/data/hces/mwi/ihs2016/hh_mod_g_clean.csv")
```

