---
title: "ihs4_cleaning"
author: "Kevin Tang"
date: "3/3/2020"
output: html_document
---

INTRO 

1) We need to identify where the data on household consumption is stored. 

For ish4, the data on food consumption was recorded as part of the household questionnaire (datasets starting w/ HH_). Within this questionnaire, data on food consumption is recorded in the module G (Food consumption). Within the module G, we find 3 that set, each one with different level of aggregation. We are interested in g1 which is food recorded at food item level (disaggregated data). 

Hence, we need to load the file hh_mod_g1 (household questionnaire module G, 1- consumption at item level). Once it is loaded, we can check that the file is loaded correctly by checking that the observations (cases) and variables match the one reported in the worldbank website (https://microdata.worldbank.org/index.php/catalog/2936/data-dictionary).



```{r}
#--- Load libraries - reminder to always load tidyverse last
library(epiDisplay)
library(plyr)
library(magrittr)
library(foreign)
library(psych)
library(readxl)
library(tidyverse)

#--- Change to show no scientific notation & round to 3 decimal places
options(scipen = 10, digits=3) 

#--- Set the plot theme to come out in black and white
theme_set(theme_bw())

#--- Read in the file

ihs4 <- read.csv(here::here("data", "hh_mod_g1.csv"))
fcode <- read.csv(here::here("data", "mwi_item_code.csv"),  fileEncoding="UTF-8-BOM")
ucode <- read.csv(here::here("data", "mwi_unit_code.csv"), fileEncoding="UTF-8-BOM")

```

## CHANGE VARIABLE NAMES AND DATA CLASS

The next step, it is to standardize the variables. This is important because we are going to be using data from different countries and the variable names will changes accross quesitonnaire. However, if we want to standardize (or harmonize) our cleaning and processing structure, it is best to standardize variables names (when possible). Within the data dictionary we can find a description of each variable. And, below is the codebook with the new variable names.

Those variables with hh_g03 refer to total consumption, hh_g04 refers to purchased by the household, hh_06 refers to food from own-production, 
hh_g07 refers to food consumed from gift (and other sources). 

Then, to identify variables that report quantity will be those that has an 'a' after the number (i.e. hh_g03a), variables reporting units will be identified by 'b' (i.e. hh_g07b, or hhg07b_label, hg07b_oth) and 'c' is for sub-unit related to photo aid guidance. 

Codebook (https://microdata.worldbank.org/index.php/catalog/2939/datafile/F192)



id =            case_id              purc_quant =    hh_g04a              gift_quant =    hh_g07a
consYN =        hh_g01               purc_unitA =    hh_g04b              gift_unitA =    hh_g07b
item_code =     hh_g02               purc_unit =     hh_g04b_label        gift_unit =     hh_g07b_label
item_oth =      hh_g01_oth           purc_unit_oth = hh_g04b_oth          gift_unit_oth = hh_g07b_oth
                                     purc_unit_size= hh_g04c              gift_unit_size= hh_g07c
                                     purc_price =    hh_g05
             
cons_quant =    hh_g03a              prod_quant =    hh_g06a
cons_unitA =    hh_g03b              prod_unitA =    hh_g06b
cons_unit =     hh_g03b_label        prod_unit =     hh_g06b_label
cons_unit_oth = hh_g03b_oth          prod_unit_oth = hh_g06b_oth
cons_unit_size= hh_g03c              prod_unit_size= hh_g06c

```{r}

#Summary info
names(ihs4)[names(ihs4) == 'case_id'] <- 'id'
names(ihs4)[names(ihs4) == 'hh_g01'] <- 'consYN'
names(ihs4)[names(ihs4) == 'hh_g02'] <- 'item_code'
names(ihs4)[names(ihs4) == 'hh_g01_oth'] <- 'item_oth'

ihs4$id <- as.character(ihs4$id)
ihs4$consYN <- as.factor(ihs4$consYN)
ihs4$item_code <- as.factor(ihs4$item_code)
ihs4$item_oth <- as.factor(ihs4$item_oth)

#G03: Quantity and units consumed of each food 
names(ihs4)[names(ihs4) == 'hh_g03a'] <- 'cons_quant'
names(ihs4)[names(ihs4) == 'hh_g03b'] <- 'cons_unitA'
names(ihs4)[names(ihs4) == 'hh_g03b_label'] <- 'cons_unit'
names(ihs4)[names(ihs4) == 'hh_g03b_oth'] <- 'cons_unit_oth'
names(ihs4)[names(ihs4) == 'hh_g03c'] <- 'cons_unit_size'

ihs4$cons_quant <- as.numeric(ihs4$cons_quant)
ihs4$cons_unitA <- as.factor(ihs4$cons_unitA)
ihs4$cons_unit <- as.factor(ihs4$cons_unit)
ihs4$cons_unit_oth <- as.factor(ihs4$cons_unit_oth)
ihs4$cons_unit_size <- as.factor(ihs4$cons_unit_size)


#G04: Quantity and units purchased of each food 
names(ihs4)[names(ihs4) == 'hh_g04a'] <- 'purc_quant'
names(ihs4)[names(ihs4) == 'hh_g04b'] <- 'purc_unitA'
names(ihs4)[names(ihs4) == 'hh_g04b_label'] <- 'purc_unit'
names(ihs4)[names(ihs4) == 'hh_g04b_oth'] <- 'purc_unit_oth'
names(ihs4)[names(ihs4) == 'hh_g04c'] <- 'purc_unit_size'
names(ihs4)[names(ihs4) == 'hh_g05'] <- 'purc_price'

ihs4$purc_quant <- as.numeric(ihs4$purc_quant)
ihs4$purc_unitA <- as.factor(ihs4$purc_unitA)
ihs4$purc_unit <- as.factor(ihs4$purc_unit)
ihs4$purc_unit_oth <- as.factor(ihs4$purc_unit_oth)
ihs4$purc_unit_size <- as.factor(ihs4$purc_unit_size)
ihs4$purc_price <- as.numeric(ihs4$purc_price)

#G06: Quantity and units home produced of each food 
names(ihs4)[names(ihs4) == 'hh_g06a'] <- 'prod_quant'
names(ihs4)[names(ihs4) == 'hh_g06b'] <- 'prod_unitA'
names(ihs4)[names(ihs4) == 'hh_g06b_label'] <- 'prod_unit'
names(ihs4)[names(ihs4) == 'hh_g06b_oth'] <- 'prod_unit_oth'
names(ihs4)[names(ihs4) == 'hh_g06c'] <- 'prod_unit_size'

ihs4$prod_quant <- as.numeric(ihs4$prod_quant)
ihs4$prod_unitA <- as.factor(ihs4$prod_unitA)
ihs4$prod_unit <- as.factor(ihs4$prod_unit)
ihs4$prod_unit_oth <-as.factor(ihs4$prod_unit_oth)
ihs4$prod_unit_size <- as.factor(ihs4$prod_unit_size)

#G07: Quantity and units gifted/other source of each food 
names(ihs4)[names(ihs4) == 'hh_g07a'] <- 'gift_quant'
names(ihs4)[names(ihs4) == 'hh_g07b'] <- 'gift_unitA'
names(ihs4)[names(ihs4) == 'hh_g07b_label'] <- 'gift_unit'
names(ihs4)[names(ihs4) == 'hh_g07b_oth'] <- 'gift_unit_oth'
names(ihs4)[names(ihs4) == 'hh_g07c'] <- 'gift_unit_size'

ihs4$gift_quant <- as.numeric(ihs4$gift_quant)
ihs4$gift_unitA <- as.factor(ihs4$gift_unitA)
ihs4$gift_unit <- as.factor(ihs4$gift_unit)
ihs4$gift_unit_oth <- as.factor(ihs4$gift_unit_oth)
ihs4$gift_unit_size <- as.factor(ihs4$gift_unit_size)
```

## MERGE IN UNIT AND ITEM LABELS

We are merging the food code with the food description and the unit with the unit description. We are also checking that the items and codes are matching. 

```{r}
ihs4 <- merge(x=ihs4, y=fcode, by.x='item_code', by.y='code', fill=-9999, all.x = TRUE)
ihs4 <- merge(x=ihs4, y=ucode, by.x='cons_unit', by.y='code', fill=-9999, all.x = TRUE)
ihs4 <- ihs4 %>% arrange(cons_unit) %>% arrange(item_code) %>% arrange(HHID)
ihs4 <- ihs4 %>% select(id:consYN, item_code, item, item_oth:cons_unitA, cons_unit, unit, cons_unit_oth:gift_unit_size)

ihs4 %>%  dplyr::count(item_code, item) %>% arrange(item_code)

ihs4 %>%  filter(item_code == "102") %>% distinct(item_code, item)

```


>CLEANING UNITS LABELLED AS "OTHER"

Food items and units are coming from a standard list. However, some food consumed by the household is not in that list of items, hence it can be recorded under 'other'. The code to identify 'other food items is

Similarly, for units, when the household is reporting the food consumed in a unit that it is not in the list will be recoded and the unit code apply for 'other' is 23.

There are a number of item units recorded as "23" but described as existing ucodes in our codebook.
We need to change these units from "23" to their appropriate units as in this analysis any item without a standard unit will be excluded. 
```{r}
ihs4 %>% filter(cons_unitA==23)
ihs4 %>% filter(cons_unit_oth!= "") %>% count(cons_unit_oth) %>% arrange(desc(n)) #which units are listed as "other"
ihs4 %>% filter(cons_unitA==23) %>% count(cons_unit_oth) %>% arrange(desc(n)) #which units are listed as "other" have not been converted to standard unit

ihs4 %>% filter(., cons_unit_oth=="HEAP") %>% select(item, item_code, cons_quant, cons_unitA, cons_unit, cons_unit_oth) %>% arrange(item)
ihs4 %>% filter(item_code=="803") %>% select(item, item_code, cons_quant, cons_unitA, cons_unit, cons_unit_oth) %>% arrange(desc(cons_unitA))

```

```{r}
ihs4[ihs4$cons_unit_oth == "HEAP", "cons_unitA"] <- 10
ihs4[ihs4$cons_unit_oth == "GRAMS", "cons_unitA"] <- 18
ihs4[ihs4$cons_unit_oth == "KILOGRAM", "cons_unitA"] <- 1
ihs4[ihs4$cons_unit_oth == "PAIL", "cons_unitA"] <- 4
ihs4[ihs4$cons_unit_oth == "TUBE /SACHET", "cons_unitA"] <- 22
ihs4[ihs4$cons_unit_oth == "SATCHET (MEDIUM)", "cons_unitA"] <- 42
ihs4[ihs4$cons_unit_oth == "HEAP (SMALL)", "cons_unitA"] <- 10
ihs4[ihs4$cons_unit_oth == "BUNDLE", "cons_unitA"] <- 8
ihs4[ihs4$cons_unit_oth == "HEAP (LARGE)", "cons_unitA"] <- "10C"
ihs4[ihs4$cons_unit_oth == "TABLESPOON", "cons_unitA"] <- 50
ihs4[ihs4$cons_unit_oth == "5L BUCKET", "cons_unitA"] <- 26
ihs4[ihs4$cons_unit_oth == "TUBE (MEDIUM)", "cons_unitA"] <- 42
ihs4[ihs4$cons_unit_oth == "PAIL (SMALL)", "cons_unitA"] <- 4
ihs4[ihs4$cons_unit_oth == "TEASPOON", "cons_unitA"] <- 20
ihs4[ihs4$cons_unit_oth == "HEAP (MEDIUM)", "cons_unitA"] <- 10
ihs4[ihs4$cons_unit_oth == "PACKET 100 GRAM", "cons_unitA"] <- 34
ihs4[ihs4$cons_unit_oth == "PAIL (LARGE)", "cons_unitA"] <- 4
ihs4[ihs4$cons_unit_oth == "TUBE (LARGE)", "cons_unitA"] <- 43
ihs4[ihs4$cons_unit_oth == "PACKET 150 GRAM", "cons_unitA"] <- 34
ihs4[ihs4$cons_unit_oth == "PIECE (SMALL)", "cons_unitA"] <- "9A"
ihs4[ihs4$cons_unit_oth == "PIECES", "cons_unitA"] <- 9
ihs4[ihs4$cons_unit_oth == "NO. 12 PLATE", "cons_unitA"] <- 7
ihs4[ihs4$cons_unit_oth == "SATCHET (LARGE)", "cons_unitA"] <- 43
ihs4[ihs4$cons_unit_oth == "PIECE MEDIUM", "cons_unitA"] <- "9B"
ihs4[ihs4$cons_unit_oth == "PIECE", "cons_unitA"] <- 9
ihs4[ihs4$cons_unit_oth == "NO. 10 PLATE", "cons_unitA"] <- 6
ihs4[ihs4$cons_unit_oth == "5 LITRES", "cons_unitA"] <- 26
ihs4[ihs4$cons_unit_oth == "PAIL (MEDIUM)", "cons_unitA"] <- "4B"
ihs4[ihs4$cons_unit_oth == "PIECE (LARGE)", "cons_unitA"] <- "9C"
ihs4[ihs4$cons_unit_oth == "MANGO", "cons_unitA"] <- 9
ihs4[ihs4$cons_unit_oth == "PACKET (SMALL)", "cons_unitA"] <- 54
ihs4[ihs4$cons_unit_oth == "SATCHET (250 GRAM)", "cons_unitA"] <- 65
ihs4[ihs4$cons_unit_oth == "SATCHET (20 GRAM)", "cons_unitA"] <- 41
ihs4[ihs4$cons_unit_oth == "TIN", "cons_unitA"] <- 72
ihs4[ihs4$cons_unit_oth == "PIECE (MEDIUM)", "cons_unitA"] <- "9B"
ihs4[ihs4$cons_unit_oth == "PACKET 250 GRAM", "cons_unitA"] <- 65
ihs4[ihs4$cons_unit_oth == "PACKETS", "cons_unitA"] <- 60
ihs4[ihs4$cons_unit_oth == "TIN (500G)", "cons_unitA"] <- 73
ihs4[ihs4$cons_unit_oth == "SACHET", "cons_unitA"] <- 22
ihs4[ihs4$cons_unit_oth == "SATCHET (MEDIUM)", "cons_unitA"] <- 22
ihs4[ihs4$cons_unit_oth == "SATCHET (500 GRAM)", "cons_unitA"] <- 22
ihs4[ihs4$cons_unit_oth == "SATCHET (SMALL)", "cons_unitA"] <- 22
ihs4[ihs4$cons_unit_oth == "SATCHET (150 GRAM)", "cons_unitA"] <- 43
ihs4[ihs4$cons_unit_oth == "BOTTLE" & ihs4$item_code == "803", "cons_unitA"] <- 15

ihs4[ihs4$cons_unit_oth == "HEAP", "cons_unit"] <- 10
ihs4[ihs4$cons_unit_oth == "GRAMS", "cons_unit"] <- 18
ihs4[ihs4$cons_unit_oth == "KILOGRAM", "cons_unit"] <- 1
ihs4[ihs4$cons_unit_oth == "PAIL", "cons_unit"] <- 4
ihs4[ihs4$cons_unit_oth == "TUBE /SACHET", "cons_unit"] <- 22
ihs4[ihs4$cons_unit_oth == "SATCHET (MEDIUM)", "cons_unit"] <- 42
ihs4[ihs4$cons_unit_oth == "HEAP (SMALL)", "cons_unit"] <- 10
ihs4[ihs4$cons_unit_oth == "BUNDLE", "cons_unit"] <- 8
ihs4[ihs4$cons_unit_oth == "HEAP (LARGE)", "cons_unit"] <- 10
ihs4[ihs4$cons_unit_oth == "TABLESPOON", "cons_unit"] <- 50
ihs4[ihs4$cons_unit_oth == "5L BUCKET", "cons_unit"] <- 26
ihs4[ihs4$cons_unit_oth == "TUBE (MEDIUM)", "cons_unit"] <- 42
ihs4[ihs4$cons_unit_oth == "PAIL (SMALL)", "cons_unit"] <- 4
ihs4[ihs4$cons_unit_oth == "TEASPOON", "cons_unit"] <- 20
ihs4[ihs4$cons_unit_oth == "HEAP (MEDIUM)", "cons_unit"] <- 10
ihs4[ihs4$cons_unit_oth == "PACKET 100 GRAM", "cons_unit"] <- 34
ihs4[ihs4$cons_unit_oth == "PAIL (LARGE)", "cons_unit"] <- 4
ihs4[ihs4$cons_unit_oth == "TUBE (LARGE)", "cons_unit"] <- 43
ihs4[ihs4$cons_unit_oth == "PACKET 150 GRAM", "cons_unit"] <- 34
ihs4[ihs4$cons_unit_oth == "PIECE (SMALL)", "cons_unit"] <- 9
ihs4[ihs4$cons_unit_oth == "PIECES", "cons_unit"] <- 9
ihs4[ihs4$cons_unit_oth == "NO. 12 PLATE", "cons_unit"] <- 7
ihs4[ihs4$cons_unit_oth == "SATCHET (LARGE)", "cons_unit"] <- 43
ihs4[ihs4$cons_unit_oth == "PIECE MEDIUM", "cons_unit"] <- 9
ihs4[ihs4$cons_unit_oth == "PIECE", "cons_unit"] <- 9
ihs4[ihs4$cons_unit_oth == "NO. 10 PLATE", "cons_unit"] <- 6
ihs4[ihs4$cons_unit_oth == "5 LITRES", "cons_unit"] <- 26
ihs4[ihs4$cons_unit_oth == "PAIL (MEDIUM)", "cons_unit"] <- 4
ihs4[ihs4$cons_unit_oth == "PIECE (LARGE)", "cons_unit"] <- 9
ihs4[ihs4$cons_unit_oth == "MANGO", "cons_unit"] <- 9
ihs4[ihs4$cons_unit_oth == "PACKET (SMALL)", "cons_unit"] <- 54
ihs4[ihs4$cons_unit_oth == "SATCHET (250 GRAM)", "cons_unit"] <- 65
ihs4[ihs4$cons_unit_oth == "SATCHET (20 GRAM)", "cons_unit"] <- 41
ihs4[ihs4$cons_unit_oth == "TIN", "cons_unit"] <- 72
ihs4[ihs4$cons_unit_oth == "PIECE (MEDIUM)", "cons_unit"] <- 9
ihs4[ihs4$cons_unit_oth == "PACKET 250 GRAM", "cons_unit"] <- 65
ihs4[ihs4$cons_unit_oth == "PACKETS", "cons_unit"] <- 60
ihs4[ihs4$cons_unit_oth == "TIN (500G)", "cons_unit"] <- 73
ihs4[ihs4$cons_unit_oth == "SACHET", "cons_unit"] <- 22
ihs4[ihs4$cons_unit_oth == "SATCHET (MEDIUM)", "cons_unit"] <- 22
ihs4[ihs4$cons_unit_oth == "SATCHET (500 GRAM)", "cons_unit"] <- 22
ihs4[ihs4$cons_unit_oth == "SATCHET (SMALL)", "cons_unit"] <- 22
ihs4[ihs4$cons_unit_oth == "SATCHET (150 GRAM)", "cons_unit"] <- 43
ihs4[ihs4$cons_unit_oth == "BOTTLE" & ihs4$item_code == "803", "cons_unit"] <- 15
```

# CLEANING FOODS LABELLED AS "OTHER"



```{r}
item_other <- ihs4 %>% count(item_oth) %>% arrange(desc(n))

ihs4 %>% filter(item_oth=="BIRD")
```

```{r}
ihs4$item_code[ihs4$item_oth=='BAKED AFRICAN CAKE'] <- 836
ihs4$item_code[ihs4$item_oth=='GRAIN MEAL'] <- 101
ihs4$item_code[ihs4$item_oth=='GRANMILL'] <- 101
ihs4$item_code[ihs4$item_oth=='MANDASI'] <- 827
ihs4$item_code[ihs4$item_oth=='PORK'] <- 506
ihs4$item_code[ihs4$item_oth=='PEANUT'] <- 304  
ihs4$item_code[ihs4$item_oth=='PIGEON PEAS'] <- 303  
ihs4$item_code[ihs4$item_oth=='CUCUMBER'] <- 409  
ihs4$item_code[ihs4$item_oth=='SMALL FISH'] <- 5023   
ihs4$item_code[ihs4$item_oth=='SODA'] <- 907 
ihs4$item_code[ihs4$item_oth=='FRIED MIXTURE OF MAIZE FLOUR/BANANA AND SODA'] <- 827 
ihs4$item_code[ihs4$item_oth=='MIXTURE OF MAIZE FLOUR/BANANA/SUGAR'] <- 827 
ihs4$item_code[ihs4$item_oth=='IRISH POTATOES AND FLOUR'] <- 821
ihs4$item_code[ihs4$item_oth=='SCONES'] <- 112 
ihs4$item_code[ihs4$item_oth=='WINE'] <- 914
ihs4$item_code[ihs4$item_oth=='TEA'] <- 901
ihs4$item_code[ihs4$item_oth=='BIRD'] <- 509
ihs4$item_code[ihs4$item_oth=='DUCKMEAT'] <- 509
ihs4$item_code[ihs4$item_oth=='COW PEAS'] <- 308
ihs4$item_code[ihs4$item_oth=='BONONGWE'] <- 403
ihs4$item_code[ihs4$item_oth=='CASSAVA LEAVES'] <- 406
ihs4$item_code[ihs4$item_oth=='CHIKONDAMOYO'] <- 836
ihs4$item_code[ihs4$item_oth=='CHISOSOANDBONONGWE'] <- 403
ihs4$item_code[ihs4$item_oth=='FRESHBEANS'] <- 302
ihs4$item_code[ihs4$item_oth=='GROUND AND COOKED BEANS'] <- 302
ihs4$item_code[ihs4$item_oth=='MAIZEMEAL'] <- 101
ihs4$item_code[ihs4$item_oth=='SOYA'] <- 314

```

## MISC. CLEANING FUNCTIONS

Create unique ID variable
```{r}
ihs4$measure_id <- paste0(as.character(ihs4$item_code),"_", as.character(ihs4$cons_unitA))
ihs4$measure_id[is.na(ihs4$cons_unit)] <- NA

ihs4 %>%  dplyr::count(item_code, item, item_oth) %>% arrange(item_code)

write.csv(ihs4, here::here("data","inter-output", "hh_mod_g_clean.csv"))

```

